---
title: "HW#1"
author: "Lucas Villanti"
date: "2024-01-26"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(corrplot)
```

```{r}
ioct <- read.csv("IOCT.csv")
```

#DATA EXPLORATION 
#Consider plots that show distributions of variables of interest. Issues of collinearity. Modifications of covariates if necessary. Too much detail will definitely cause DPE to lose interest, too little detail will result in questions for your final models.

```{r}
#Histogram for Height
ioct %>%
  select(height) %>%
  ggplot(aes(x = height)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Height Distribution", x = "Height", y = "Count")
```

```{r}
#Histogram for Weight
ioct %>%
  select(weight) %>%
  ggplot(aes(x = weight)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Weight Distribution", x = "Weight", y = "Count")
```

```{r}
#Histogram for IOCT Times
ioct %>%
  select(IOCT_Time) %>%
  ggplot(aes(x = IOCT_Time)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "IOCT time Distribution", x = "IOCT TIMES", y = "Count")
```

```{r}
#Correlation matrix

cor_matrix <- cor(ioct[, c("height", "weight", "IOCT_Time", "pu_score", "su_score", "run_score", "APFT_Score")])
corrplot(cor_matrix, method = "circle")
```

#DATA PREPARATION - Consider outliers. What will you do with highpoints that have high leverage? Will you transform any of your variables?

```{r}
#Looking for outliers

#scatter plot for ioct and height
plot(ioct$height,ioct$IOCT_Time,main = "IOCT vs Height",xlab = "Height", ylab = "IOCT_Time")

#Going to remove the highest IOCT time(which is the outlier)
ioct <- ioct[ioct$IOCT_Time < max(ioct$IOCT_Time), ]
```


```{r}
#I will seperate the females and Males because there are different standards
female_ioct <- ioct %>%
  filter(sex == "F")
male_ioct <- ioct %>%
  filter(sex == "M")
```

#BUILD MODELS - Consider at least 3 different models, using different variables (or different transformations). If your models are nested conduct formal tests to decide on which model is the most appropriate. If the models are not nested consider using AIC or BIC to decide
```{r}
#Model 1 Height for Male
model1_male <- lm(IOCT_Time ~ height, data = male_ioct)
summary(model1_male)

#Model 1 Height for Female
model1_female <- lm(IOCT_Time ~ height, data = female_ioct)
summary(model1_female)

#model 1 height for all
model1 <- lm(IOCT_Time ~ height, data = ioct)
summary(model1)
```

```{r}
#Model 2 with everything Male
model2_male <- lm(IOCT_Time ~ height + weight + pu_score + su_score + run_score, data = male_ioct)
summary(model2_male)

#Model 2 with everything Female
model2_female <- lm(IOCT_Time ~ height + weight + pu_score + su_score + run_score, data = female_ioct)
summary(model2_female)

#Model 2 not seperating
model2 <- model2 <- lm(IOCT_Time ~ height + weight + APFT_Score + sex, data = ioct)
summary(model2)
```

```{r}
#model 3 - interaction : Male
model_interaction_male <- lm(IOCT_Time ~ height + weight + pu_score * height + su_score + run_score, data = male_ioct)
summary(model_interaction_male)

#model 3 - interaction : Female
model_interaction_female <- lm(IOCT_Time ~ height + weight + pu_score * height + su_score + run_score, data = female_ioct)
summary(model_interaction_female)

#model 3 not seperating male or female
model3 <- lm(IOCT_Time ~ height * weight + APFT_Score + sex + CS, data = ioct)
summary(model3)
```


```{r}
#Testing Model 2 & 3 with anova
anova(model2,model3)
```

```{r}
#testing using Log Likelyhood test
library(lmtest)

lrtest(model2,model3)
```


#Final Model - Final model - Present your best model. Why did you select it? What is your answer to the research question? What are some limitations for your study? Based off of your model what is your predicted IOCT time?

```{r}
summary(model2)
predict(model2, newdata = data.frame(height = 73, weight = 200, APFT_Score = 257, sex = "M"))
```

